---
title: "AMR Summary Statistics"
author: "Brooke Watson"
date: "November 14, 2016"
output: html_document
---



## AMR Summary Statistics 

```{r, include=FALSE} 
knitr::opts_chunk$set(echo = TRUE)
```

### 1. Setup: Read in the munged datasets.

```{r message=FALSE, results='hide', warning=FALSE, message=FALSE}
rm(list=ls())
library(knitr)
knit('EIDR_cleaning_home.Rmd', tangle=TRUE)
source('EIDR_cleaning_home.R')
amr = filter(all, pathogenDrugResistanceVal == "yes")
write.xlsx(amr, "amr_refs.xlsx", sheetName="AMR References from EIDR") 
# length(unique(amr$Title))

```

### 2. Other subsets to match with the EMBASE data.

For the AMR research we want to look just at bacteria and refine our focus to the 1970-2016 time frame, so let's filter this further. 

```{r message=FALSE, results='hide', warning=FALSE, message=FALSE}

bac = filter(amr, pathogenTypeVal == 'bacteria')
write.xlsx(bac, "amr_refs.xlsx", append = TRUE, sheetName="Restricted to Bacteria") 

year_filtered = filter(bac, year>=1970)
write.xlsx(year_filtered, "amr_refs.xlsx", append = TRUE, sheetName="Real Filtered by Year") 

```

### 3. Getting Summary Statistics. 

Here we're going to convert everything to numeric and look at distributions. 
Later we'll do the same thing with factors - this is just to get at what are the actual numeric values in the dataset, and what are the distributions of their missing values. 


 ### Frequency of events by country in total EIDR database.  
    
```{r Frequency of events by country - All EIDR} 

tab = as.numeric(table(all$locationNationVal))
names(tab) = names(table(all$locationNationVal))
tab = data.frame(Country = names(tab), N_events = tab)
tab = arrange(tab, desc(N_events))
tab$Country <- factor(tab$Country, levels = tab$Country[order(tab$N_events)])

kable(tab, booktabs = TRUE)
 
p = ggplot(arrange(tab, desc(N_events)), aes(Country, N_events)) +
           geom_bar(stat = "identity") +    
    ggtitle("All Countries with an emergence event listed in EIDR") +
           theme(axis.text.y = element_text(angle = 0, hjust = 1)) 
p + coord_flip()

tab2 = filter(tab, N_events>1)
p = ggplot(arrange(tab2, desc(N_events)), aes(Country, N_events)) +
           geom_bar(stat = "identity") + 
    ggtitle("Countries with more than one emergence event") +
           theme(axis.text.y = element_text(angle = 0, hjust = 1)) 
p + coord_flip()
 

```          


### Frequency of events by country, restricted to bacterial antibiotic resistance events.  
    
```{r Frequency of events by country - Bacterial, Resistant} 


tab = as.numeric(table(bac$locationNationVal))
names(tab) = names(table(bac$locationNationVal))
tab = data.frame(Country = names(tab), N_events = tab)
tab = arrange(tab, desc(N_events))
tab$Country <- factor(tab$Country, levels = tab$Country[order(tab$N_events)])
tab$Country <- factor(tab$Country, levels = tab$Country[order(tab$N_events)])


p = ggplot(tab, aes(Country, N_events)) +
           geom_bar(stat = "identity") +
           theme(axis.text.x = element_text(angle = 0, hjust = 1)) 
p + coord_flip()
           
```           