---
title: "AMR Summary Statistics"
output: html_notebook
---

## AMR Summary Statistics 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Setup: Read in the munged datasets.

```{r message=FALSE, results='hide', warning=FALSE, message=FALSE}
rm(list=ls())
knit('EIDR_cleaning_home.Rmd', tangle=TRUE)
source('EIDR_cleaning_home.R')
amr = filter(all, pathogenDrugResistanceVal == "yes")
write.csv(amr, file = "amr_refs.csv")
# length(unique(amr$Title))

```

# 2. Getting Summary Statistics. 

Here we're going to convert everything to numeric and look at distributions. 
Later we'll do the same thing with factors - this is just to get at what are the actual numeric values in the dataset, and what are the distributions of their missing values. 

```{r Summary Statistics}
  
lapply(all, as.numeric) 
sumlist = lapply(allnum, summary)
 

row1 = summary(allnum$eidCategoryVal)
row2 = summary(allnum$eventName)
sumstats = rbind(row1, row2)
allnumshort = rbind(as.character(sumlist))
summarydf = lapply(allnumshort, strsplit, split = ",")
 
```

NAlist = lapply(NAlist, unlist)
test = nas(all, 1)
select

# summary stats for the numeric variables
nums = lapply(all, as.numeric)
lapply(nums, summary) # so most things are character vectors 

test= lapply((1:ncol(all)), nas, db = all)
test= lapply((1:ncol(nums)), nas, db = nums)
nums = nums[ , (sum(is.na(nums)) < 200)]
nums = nums %>% as.data.frame()

 
``` 
 
# Plots 
    
```{r Plots}    
    
qplot(as.numeric(as.factor(all$pathogenTypeVal)), xlab = names(as.factor(all$pathogenTypeVal))
          
```           