---
title: "EIDR summary"
author: "Brooke"
date: "11/8/2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Attaching Necessary Packages.

```{r Attaching Necessary Packages, warning=FALSE, message=FALSE, results='hide'}

pkgs = list('knitr', 'readxl', 'dplyr', 'xtable', 'stringr', 'devtools', 'plotly', 'janitor', 'stringi', 'xlsx', 'bookdown')
lapply(pkgs, library, character.only = TRUE)
 
```

# 2. Importing Data.  

Note - don't forget to change the working directory.

## a.) Clear environment, set working directory, and import xlsx file using readxl.

```{r Setup Working Directory}
rm(list=ls()) 
wd = "~/Dropbox (EHA)/GVP/EIDR/"
wd2 = '~/Desktop/git/eidr' # for when using a different computer 
setwd(wd)
path = "~/Dropbox (EHA)/AMR/data/EIDR Spreadsheet.xlsx"
path2 = paste0(getwd(), "~/Desktop/EIDR Spreadsheet.xlsx")
list = excel_sheets(path) # see names of excel sheets 
```

## b.) Read in the source code that will assign multiple objects in the global environment at once using an lapply loop.  

```{r Source Github function, warning=FALSE, message=FALSE}

source('https://gist.githubusercontent.com/brooke-watson/e03c8a785fe69cbef42a7ebfe9e24613/raw/5782d93be6b49307988a211231ee5f309298cf63/multassign.R') # find the function 
g(event, pathogen, host, location, col_names) %=%
  lapply(excel_sheets(path)[c(1:4, 6)], read_excel, path=path)   # loading specific excel sheets 
```

# 3. Cleaning Data. 

## a.) Filter out non-EIDs and merge the datasets. 

```{r Merge and Clean Datasets}

event = filter(event, eidVal == 1)

# merge data
all = merge(event, host, by='eventName') 
all = merge(all, pathogen, by='eventName')
all = merge(all, location, by='eventName')
``` 
## b.) select variables and make a duplicate backup dataset just in case 

```{r}

all = all %>%
  dplyr::select(eventName, eidCategoryVal, eidVal, numberInfectedVal, driverVal,
                diseaseVal, startDateISOVal, numberDeathsVal, zoonoticVal, testingMethodVal,
                pathogenTypeVal, pathogenFamilyVal, pathogenDrugResistanceVal,
                pathogenOrderVal, DomesticationStatusVal, hostVal, refAbstract, 
                hostClassVal, hostFamilyVal, hostSpeciesVal, hostTaxOrderVal,
                locationNationVal, locationContinentVal,
                refEventName, refEidCategory, refAbstract, refDriver, refDisease, 
                refStartDateISO) %>% unique()
# unique(all$pathogenTypeVal)  

```


## c.) Character Munging - trimming white space, making everything lower case, dropping duplicates. 

```{r Character Munging }

all = lapply(all, trimws) %>% # trim white space
    lapply(tolower) %>% #standardize capitalization 
    as.data.frame(stringsAsFactors = FALSE) %>% # turn back into a data frame 
    unique() # drop duplicates 
all2 = all #making a backup just in case 

``` 


## d.) Comparing and fixing references. 

Before moving forward, we need to make sure all the events have a reference, as we're going to join those to a Zotero table later. The reference documentation is a little iffy, though, so we'll compare the difference reference columns and choose one that is most complete. 

```{r Ref fix}

refdb = all %>% 
    dplyr::select(refEventName, refEidCategory, refAbstract, refDriver, refDisease, 
                refStartDateISO) # creating a reference-specific database to compare                                         # differences. 

# get number of NAs per reference column to see which is most complete 
nums = lapply(refdb, is.na) %>% 
    lapply(sum) %>% rbind() %>% print() 
```

refEidCategory is the most complete reference variable (Not counting refEventName, which references the reviews and not the individual event reports and therefore can't be used for this analysis.) To get one reference per event, do the following: 

1. See if observations that are NA in the most complete list (refEidCategory) have a reference listed in a different ref column. 

```{r}
 NAdb = filter(refdb, is.na(refEidCategory)) # checking those with no references 
 # View(NAdb)
```

2. Fill in NA's in EidCategory with references from the other ref columns. 

```{r, message=FALSE, warning=FALSE}
if (is.na(refdb$refEidCategory) & !is.na(refdb$refAbstract)){
    refdb$refEidCategory = refdb$refAbstract
} else {refdb$refEidCategory = refdb$refEidCategory}
```

3.. Extract just the first reference from the most complete list, and add it to the "all" database. 

```{r}
all = dplyr::select(all, -c(refEidCategory, refAbstract, refDriver, refDisease, refStartDateISO)) # remove reference columns 

all$ref = gsub( ",.*$", "", refdb$refEidCategory) %>% as.integer() 
# add single numeric reference column 
rm(NAdb)

```

join up reference codes to reference spreadsheet 

```{r, warning=FALSE, message=FALSE}

refs = read_excel(paste0(getwd(), "/EIDR References.xlsx"), sheet = 1) %>% 
    janitor::clean_names()
# install_github('brooke-watson/thesis') 
# library)
refs = select(refs, c(rights, author, title, publication_title, doi, issn, url, access_date)) %>% 
    dplyr::rename(ref = rights) %>% 
    lapply(trimws) %>% # trim white space
    lapply(tolower) %>% #standardize capitalization 
    as.data.frame(stringsAsFactors = FALSE)     

all = merge(all, refs, by="ref")

```
 
This dataset still has 362 observations, while online EIDR has 350. Checking the duplicates using the `janitor` package, we can see that they're all caused by different location values.

```{r Checking Duplicates}
dupes = get_dupes(all, eventName)  
# View(dupes) 
```  

# That's fine. We'll keep those for now. Moving on. 

##. d.) Creating Attributes. 

Here we want to get number of unique values per column and set it as an attribute. This helps get at how many different possible categorical answers there are for string variables. 

```{r Creating Attributes}
un = lapply(all, unique) %>% sapply(length)     # int vector, n of unique values per variable
attr(all, "uniques") = un                       # setting uniques as an attribute varaiable 
summary(un)                                     # range of unique values 
names(all[attr(all, "uniques") >  109] )        # seeing at a glance which ones are huge (top quartile)
# all = all[attr(all, "uniques") <  109 | attr(all, "names") == 'eventName']          
# subsetting just the reasonable ones + uniqueID. 
```

## e. Converting Strings to Numbers.  

For the ones for which that makes sense. As a reminder, it's helpful to look at `names(all)` to see which values would even make sense as numbers. One could also just check the drop down arrow in the data subheading of the environment panel in RStudio. That's what I did, so I didn't include that code. 

Additionally, each event has different time frames for its startDateISOVal. Some have days, others months, and some only have a year. They seem to all have a year, though, so we're going to go with the least common denominator. They're also conveniently all listed year-first, so we don't have to do any complicated regex backflips. 

```{r As.Numeric(), warning=FALSE, message=FALSE, error=FALSE} 
# creating a year variable and removing the startdateiso variable from the data frame. 
all$year = lapply(all$startDateISOVal, substr, start = 1, stop = 4) 
all$year = as.numeric(all$year)
all = dplyr::select(all, -startDateISOVal)
```

## f. Pulling binary encodings out of things stuck in messy strings. 

Here we have a column variable called "hostVal" that includes character strings of hosts - both human and animal - in which the pathogen was found (or theorized to have colonized). Some have long lists of hosts, and some have only one. Some are a specific family or genus, others use common names, some are specific while others mention a group or category (e.g.) birds vs chickens vs aves vs "domestic avian species.")

```{r Head Host Val}
head(all$hostVal, 6)
```

Gross. Unhelpful. 

We're not interested in ALL of this mess, but we may want to pull some values out for summary statistics and future use in regressions and other statistical analyses. To analyze, we need encoded variables, not character vectors.

The function `dummyv` (Dummy Variable) takes care of that. it takes the following arguments: 
1.) A value - i.e., the character string you'd like to extract from a long list of character strings 
2.) A data frame 
3.) The column of the data frame where you want to look for the value. 

dummyv initializes an empty column of zeroes and tacks it onto the end of your data frame. The name of the column will be a combination of the variable and the value name, joined by an underscore like so: `ColumnName_Value`. 
dummyv then looks down the column for every instance of the value you're looking for. Every row that contains the value will get a 1 instead of the 0 in the new `ColumnName_Value` column.

You can then get frequencies of that value in your dataset, subset by that value, compare differences of other variables, and go for gold with your normal analyses. 

```{r Encoding the Data}
# create binary dummy variables from lists 
# wd is set in the "Setup Working Directory" chunk - flip between the two as necessary 
source(paste0(getwd(), '/dummyv.R'))  # change the path file 
source(paste0(getwd(), '/spellcheck.R'))  # change the path file 

       # (((( don't forget to make these a function later  ))))
       
       
all = dummyv('human', all, 'hostVal') #testing a human binary variable 
# all = dummyvh('human') # custom dummyv variable for host 


head(all$hostVal, 10)

# all = all2
#tail(all$hostVal)
# all = spellcheckh('pig', 'swine', max.dist = 1)
# tail(all$hostVal)
```

 

```{r}
# hostlist = list('human', 'primate', 'pig', 'rodent', 'bat', 'cat', 'dog', 'sheep', 'goat')
# 
#  fix this later  
```

#7. Fixing spelling 

There seem to be misspellings of actinopterygii and mammalia. Let's change that. 

```{r Fix Spelling}
mam = agrep(pattern = "mammalia", x = all$hostClassVal, ignore.case = TRUE, value = FALSE, max.distance = 3) 
act = agrep(pattern = "actinopterygii", x = all$hostClassVal, ignore.case = TRUE, value = FALSE, max.distance = 5) 
  
all[mam, "hostClassVal"] = "mammalia" 
all[act, "hostClassVal"] = "actinopterygii"  
``` 

Check the new list to make sure it worked. 

```{r, eval=FALSE}
unique(all$hostClassVal)
```

***  
