---
title: "AMR_ref_merge"
author: "Brooke Watson"
date: "November 14, 2016"
output: word_document
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Read in the amr references from EIDR. 
* tbh this should not be a long term solution. it involves knitting and sourcing the amr.Rmd file, which itself knits and sources the EIDR_cleaning.Rmd file. We're getting into an untenable Russian nesting doll sitution though.  


```{r EIDR}
wd = "/Users/Watson/Dropbox (EHA)/GVP/EIDR"
knit('amr.Rmd', tangle=TRUE)
source('amr.R')
wd = "/Users/Watson/Dropbox (EHA)/GVP/EIDR"
eidref = read_excel(paste0(wd, '/amr_refs.xlsx'))
 
```

2. set working directory, load custom packages, and source a function (tidyembase) that takes raw embase output, sets the search strings as a table attribute, and then reomves that row and cleans up the data frame. 

```{r EIDR importing, warning=FALSE, message=FALSE}
amrwd = "~/Dropbox (EHA)/AMR/Systematic Review"
install_github('brooke-watson/thesis')
library(thesis)   
source('~/Dropbox (EHA)/R-scraps/unpackaged undocumented/tidyembase.R')

```

3. Read in and merge the references from the systematic review. These had to be taken out in 3 chucks because embase can only export 10,000 rows at a time, so they're sorted by date.  

```{r}

embase1 = tidyembase(paste0(amrwd, '/records_1970-2004.xlsx'))
embase2 = tidyembase(paste0(amrwd, '/records_2005-2012.xlsx'))
embase3 = tidyembase(paste0(amrwd, '/records_2013-2016.xlsx'))


```

4. Merge and clean. 

This includes pulling the DOIs from the full text link in the embase data frame so they can be matched with EIDR.   


```{r}

embase = dplyr::bind_rows(embase1, embase2, embase3)  
 
embase = embase %>% 
    janitor::clean_names() %>% 
    munge %>% 
    select(title, author_names, source, abstract, full_text_link)  
embase$doi = lapply(embase$full_text_link, sub, 
                    pattern = "http://dx.doi.org/", replacement = "")

rm(embase1, embase2, embase3)
```

5. All clean. Now fuzzy match em. Matching on doi first, and then trying to match on title and authors just in case.

```{r}

refmatch = data.frame(eidref$doi, eidref$title, eidref$author)
refmatch$embasedoi <- "" # Creating an empty column
refmatch$embaseauthor <- ""
refmatch$embasetitle <- ""

for(i in 1:dim(refmatch)[1]) {
 x <- agrep(refmatch$eidref.doi[i], embase$doi,
 ignore.case=TRUE, value=TRUE,
 max.distance = 0.05, useBytes = TRUE)
 x <- paste0(x,"")
 refmatch$embasedoi[i] <- x

 y = agrep(refmatch$eidref.author[i], embase$author_names,
 ignore.case=TRUE, value=TRUE,
 max.distance = 0.05, useBytes = TRUE)
 y <- paste0(y,"")
 refmatch$embaseauthor[i] <- y

 z = agrep(refmatch$eidref.title[i], embase$title,
 ignore.case=TRUE, value=TRUE,
 max.distance = 0.05, useBytes = TRUE)
 z <- paste0(z,"")
 refmatch$embasetitle[i] <- z
 
}
```

4. Write to a new excel sheet. Put the number of references used in the name of the file to ID between different systematic review searches. 

```{r}
amr$embasedoi = refmatch$embasedoi 
amr$embaseauthor = refmatch$embaseauthor
amr$embasetitle = refmatch$embasetitle

n = dim(embase)[[1]]

write.xlsx(amr, paste0("amr_refs_", n, ".xlsx"), sheetName="AMR References from EIDR") 
  
```

